---
title: "Project 6 - Risk-Taking Decisions"
author: "Ye Dam Yi"
date: "3/30/2021"
output: html_document
---

# Risk-Taking Decisions and Emotions

```{r load-libraries}

library(tidyverse)

```

```{r read-data}
# import data
risk <- read_csv("/Users/yedam/Documents/Academic/WFU/Psychology/Projects/First Year Project/Main/Data/Data Files/clean data_excluded.csv", na = c("", "NA"))
glimpse(risk)
```
```{r select-data}
# drop some variables and convert column names into lowercase
risk <- risk %>% 
  select(-c("StartDate", "EndDate", "Finished")) %>%
  rename_with(tolower) %>% 
  mutate(gender = as.factor(gender))
glimpse(risk)

# check factor levels - gender, recipient, experiencedemotions, condition
nlevels(risk$gender)
nlevels(risk$recipient)
nlevels(risk$experiencedemotions)
nlevels(risk$condition)

# convert into categorical variables
risk <- risk %>% 
  mutate(recipient = as.factor(recipient),
         experiencedemotions = as.factor(experiencedemotions),
         condition = as.factor(condition))

# write data
write_csv(risk, path = "data/risk.csv")
```


### Q1: Do people more risk-taking decisions for others than for themselves?
```{r risk-taking-self-other}
# plot data
risk %>% 
  ggplot(aes(y = risktaking, color = recipient)) +
  geom_boxplot() +
  labs(
    color = "Recipient",
    y = "Risk Takng Levels",
    x = "Self vs. Other"
  ) + 
  scale_color_hue(labels = c("Self", "Other"))

# look at descriptive info
risk %>% 
  select("risktaking", "recipient") %>% 
  group_by(recipient) %>% 
  summarize(mean = mean(risktaking),
            sd = sd(risktaking),
            min = min(risktaking),
            max = max(risktaking),
            median = median(risktaking))

# test this trend
m_self_other <- lm(risktaking ~ recipient, risk)
summary(m_self_other)

```

The plot and the inferential test show that people take more risk for others (M = 2.83) than for themselves (M = 2.14), p <.001. 




### Does that trend differ for females vs. males?


### Do people making decisions for others feel more or less negative emotions? Do negative emotions explain risk-taking levels? Do negative emotions influence decision making differently for personal decisions and for surrogate decisions?


### Do people making decisions for others feel more or less positive emotions? Do positive emotions explain risk-taking levels? Do positive emotions influence decision making differently for personal decisions and for surrogate decisions?


### What's the model of risk-taking levels? Is the effect of negative and positive emotions on risk-taking levels moderated by who is being decided for?





